Modelling the emergence of whisker barrels

### 2019-12-17 paper/briefpaper/brief.tex
Fixes a mis-placed white background box in figure, also tweaks to caption

Software tools for modelling the cortical sheet as a hexagonal lattice with no-flux boundary conditions and an arbitrary boundary shape, for computing the flux using a finite difference method, for numerical integration by the fourth-order Runge-Kutta method, and for reproducing the reported results are available at \url{https://github.com/ABRG-Models/BarrelEmerge}.

@@

\caption{\textbf{A.} Cortical whisker barrels from \cite{zheng_signal_2001}
  annotated with lines joining interstices into a putative Voronoi diagram
  ($\delta=0.022$). \textbf{B.}  Top: Stochastic initial condition for $a_i$
  with $i=1$ (barrel `$\alpha$'). The roll-off towards the boundary (thin
  line) ensures the boundary conditions are satisfied. Bottom: Guidance
  molecules $\rho_1(\mathbf{x})$ and $\rho_2(\mathbf{x})$ in the simulated
  cortex. $\rho_1(\mathbf{x})$ varies along the anterior-posterior axis;
  $\rho_2(\mathbf{x})$ is expressed along the lateral/medial axis, analogous
  to ephrin A5 \citep{vanderhaeghen_mapping_2000}. \textbf{C.}  Barreloid
  centres for mystacial whiskers taken from \cite{haidarliu_size_2001} are
  shown as black circles. The dashed grey line shows the boundary of the
  barreloid field. It is assumed that there are two linearly increasing,
  orthogonal signalling molecules in the thalamus, such that the location of
  each barreloid within the graded thalamic signal field can differentiate the
  efferent axons, with the result that axons from each barreloid interact to
  varying degrees with the cortical signalling gradients $\rho_1$ and
  $\rho_2$. Blue (green) arrows show, for each barreloid, the magnitude (and
  sign) of the interaction that its thalamocortical projection will have with
  $\rho_1$ ($\rho_2$). The mappings from barreloid location to interaction
  strength are linear. \textbf{D.}  Barrel cortex simulation with parameters
  $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and
  $\delta{t}=0.0001$. Colours indicate $\arg\max_i(c_i)$ for 41 TC
  projections.  \textbf{E.}  The Honda Voronoi measure, $\delta$, (red) and a
  pattern fitting metric based on the distance, $d_i$, between simulated and
  experimental barrel centroids and the absolute difference in their areas,
  $\Delta A_i$ given by $\eta = \sum_i^N d_i^2 \times \sum_i^N \Delta A_i$
  (black, units L$^4$) plotted as a function of time for the simulations shown
  in D. \textbf{F.}  Guidance molecules in the simulated Fgf8 misexpression
  experiment. $\rho_1$ is mirrored over the anterior-posterior axis. The
  simulation indicates that the mirroring of the barrel field can occur as a
  result of the mirroring of guidance molecule 1.}
\label{fig:main}
\end{figure*}

@@

The term in parentheses in Eq.~\ref{eq:dc} represents competition between thalamocortical projections for a limited availability of cortical connections. To introduce competition also in terms of axon branching we redefined
%
\begin{equation} \label{eq:comp}
\chi_i(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{j \ne i}^{N} a_{j},
\end{equation}
%
which reduces branching for each projection where the branches of other projections are dense. Divisive normalization keeps the axon branch density bounded on each iteration to the initial density  at $t=0$;

\begin{equation} \label{eq:norm}
  a_i(\mathbf{x}, t) = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
  a'_i(\mathbf{x}, t) d\sigma},
\end{equation}
%
where prime symbols indicate the use of intermediate values computed from Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$.

>>>>

Software for modelling the cortical sheet as a hexagonal lattice with no-flux
boundary conditions and an arbitrary boundary shape, for computing the flux
using a finite difference method, for numerical integration by the
fourth-order Runge-Kutta method, and for reproducing these results is available at \url{https://github.com/ABRG-Models/BarrelEmerge}.

@@

\caption{\textbf{A.} Cortical whisker barrels from \cite{zheng_signal_2001}
  annotated with lines joining interstices into a putative Voronoi diagram
  ($\delta=0.022$). \textbf{B.}  Top: Stochastic initial condition for $a_1$
  (barrel `$\alpha$'). The roll-off towards the boundary (thin line) ensures
  the boundary conditions are satisfied. Bottom: Guidance molecules
  $\rho_1(\mathbf{x})$ and $\rho_2(\mathbf{x})$ in the simulated
  cortex. $\rho_1(\mathbf{x})$ varies along the anterior-posterior axis,
  $\rho_2(\mathbf{x})$ along the lateral/medial axis (cf.~ephrin A5,
  \cite{vanderhaeghen_mapping_2000}). \textbf{C.}  Barreloid centres for
  mystacial whiskers taken from \cite{haidarliu_size_2001} are shown as black
  circles; the dashed grey line indicates the boundary. Assuming two linearly
  increasing, orthogonal signalling molecules in the thalamus, the location of
  each barreloid within this signal field can differentiate the efferent
  axons. Axons from each barreloid interact to varying degrees with the
  cortical signals $\rho_1$ and $\rho_2$. Blue (green) arrows show, for each
  barreloid's projection, the magnitude (and sign) of the interaction that it
  will have with $\rho_1$ ($\rho_2$). \textbf{D.}  Barrel cortex simulation
  with parameters $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and
  $\delta{t}=0.0001$. Colours indicate $\arg\max_i(c_i)$ for 41 TC
  projections; black lines indicate boundaries. \textbf{E.}
  The Honda Voronoi measure, $\delta$, (red) and a pattern fitness metric
  based on the distance, $d_i$, between simulated and observed barrel
  centroids and the absolute difference in their areas, $\Delta A_i$ given by
  $\eta = \sum_i^N d_i^2 \times \sum_i^N \Delta A_i$ (black, units mm$^4$)
  plotted as a function of time for the simulations shown in D. \textbf{F.}
  Guidance molecules and resulting pattern in the simulated Fgf8 misexpression
  experiment. Parameters as in D, but $\rho_1$ is mirrored over the
  anterior-posterior axis. The simulation indicates that the mirroring of the
  barrel field can occur as a result of the mirroring of guidance molecule 1.
  All scale bars are 1~mm.}
\label{fig:main}
\end{figure*}

@@

The term in parentheses in Eq.~\ref{eq:dc} represents competition between thalamocortical projections for a limited availability of cortical connections. To introduce competition also in terms of axon branching we redefined
%
\begin{equation} \label{eq:comp}
\chi_i(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{j \ne i}^{N} a_{j},
\end{equation}
%
which reduces branching for each projection where the branches of other projections are dense. Divisive normalization keeps the axon branch density bounded on each iteration to the initial density  at $t=0$;
%
\begin{equation} \label{eq:norm}
  a_i(\mathbf{x}, t) = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
  a'_i(\mathbf{x}, t) d\sigma},
\end{equation}
%
where prime symbols indicate the use of intermediate values computed from Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$.
